---
title: "Homework Assignment 1"
author: "Seokjun Choi"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

\noindent Problem 1. 
Use the qnorm function in R to find the quartiles 
(i.e., 25th, 50th and 75th percentiles) 
of the normal distribution with mean 100 and standard deviation 10. 

```{r problem1}
qnorm(c(0.25, 0.5, 0.75), 100, 10)
```

\vspace{1cm}

\noindent Problem 2. 
Use the curve function in R to display the graph of a $\chi^2(10)$ (10 corresponds to the degrees of freedom). 
Use a range of 0 to 100 for the x-axis. 
The chi-square density function is dchisq. 

```{r problem2, echo=TRUE}
curve(dchisq(x, 10), xlim = c(0, 100))
```


\noindent Problem3. (Albert and Rizzo Chapter 1, problem 3.)
(Gamma densities). Use the curve function to display the graph of the
gamma density with shape parameter 1 and rate parameter 1. Then use the
curve function with add=TRUE to display the graphs of the gamma density
with shape parameter k and rate 1 for 2,3, all in the same graphics window.
The gamma density function is dgamma. Consult the help file ?dgamma to see
how to specify the parameters.
```{r problem3, echo=TRUE}
curve(dgamma(x, shape = 1, rate = 1), xlim = c(0, 10),
  ylim = c(0, 1), xlab = "gamma densities")
curve(dgamma(x, shape = 2, rate = 1), add = TRUE, col = "red")
curve(dgamma(x, shape = 3, rate = 1), add = TRUE, col = "blue")
legend(7, 0.99, c("gamma(1,1)", "gamma(2,1)", "gamma(3,1)"), lty = 1,
  col = c("black", "red", "blue"), cex = 1)
```

\noindent Problem4. (Albert and Rizzo Chapter 1, problem 5.)
(Binomial CDF). Let X be the number of “ones” obtained in 12 rolls
of a fair die. Then X has a Binomial(n = 12, p = 1/3) distribution. 
Compute a table of cumulative binomial probabilities (the CDF) 
for x = 0,1,...,12 by two methods: 
(1) using cumsum and the result of Exercise 1.4, and 
(2) using the pbinom function. What is P(X > 7)?

Following the direction of (1), I'll use 'dbinom' function
```{r}
cumsum(dbinom(0:12, 12, 1 / 3))
```
```{r}
1 - cumsum(dbinom(0:12, 12, 1 / 3))[8]
```

Next, following (2), using 'pbinom' function,
```{r}
pbinom(0:12, 12, 1 / 3)
```
```{r}
1 - pbinom(7, 12, 1 / 3)
```
Two vectors of cdf values are exactly same.
The P(x>7) values are also same.



\noindent Problem5. (Albert and Rizzo Chapter 1, problem 7.)
(Simulated “horsekicks” data). The rpois function generates random
observations from a Poisson distribution. In Example 1.3, we compared the
deaths due to horsekicks to a Poisson distribution with mean $\lambda = 0.61$, and
in Example 1.4 we simulated random Poisson($\lambda = 0.61$) data. Use the rpois
function to simulate very large (n = 1000 and n = 10000) Poisson($\lambda = 0.61$)
random samples. Find the frequency distribution, mean and variance for the
sample. Compare the theoretical Poisson density with the sample proportions
(see Example 1.4).

When n=1000,
```{r}
pois_sample_1000 <- rpois(1000, 0.61)
empirical_1000 <-  table(pois_sample_1000) / 1000
theoretical_1000 <- dpois(0:(length(empirical_1000) - 1), 0.61)
rbind(empirical_1000,  theoretical_1000)
```
```{r}
mean(pois_sample_1000)
var(pois_sample_1000)
```

When n=10000,
```{r}
pois_sample_10000 <- rpois(10000, 0.61)
empirical_10000 <-  table(pois_sample_10000) / 10000
theoretical_10000 <- dpois(0:(length(empirical_10000) - 1), 0.61)
rbind(empirical_10000,  theoretical_10000)
```
```{r}
mean(pois_sample_10000)
var(pois_sample_10000)
```
The result of n=10000 is closer to the theoretical result than the case of n=1000.


\noindent Problem6. (Albert and Rizzo Chapter 1, problem 8.)

\noindent Problem7. (Albert and Rizzo Chapter 1, problem 9.)
\noindent problem8. (Albert and Rizzo Chapter 1, problem 10.)
\noindent problem9. (Albert and Rizzo Chapter 1, problem 11.)
\noindent problem10. (Albert and Rizzo Chapter 1, problem 12. )
\noindent problem11. (Albert and Rizzo Chapter 2, problem 3.) Briefly describe your findings.
\noindent problem12. (Albert and Rizzo Chapter 2, problems 4 and 5.) Briefly describe your findings.
\noindent problem13. (Albert and Rizzo Chapter 2, problems 7 and 8.) Briefly describe your findings.
\noindent problem14. (Albert and Rizzo Chapter 2, problems 10 and 11.)
\noindent problem15. (James et al. Chapter 2, problem 8.)
\noindent problem16. (James et al. Chapter 2, problem 10.)