---
title: "Homework Assignment 2"
author: "Seokjun Choi"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



### \noindent Problem1.
**(Albert and Rizzo Chapter 3, problem 1.)**

**(Fast food eating preference). Fifteen students in a statistics class**
**were asked to state their preference among the three restaurants Wendys,**
**McDonalds, and Subway. The responses for the students are presented below.**

Wendys McDonalds Subway Subway Subway Wendys
Wendys Subway Wendys Subway Subway Subway
Subway Subway Subway

**a. Use the scan function to read these data into the R command window.**

```{r}
# restaurant <- scan(what="character")
restaurant <- c("Wendys", "McDonalds", "Subway", "Subway", "Subway", "Wendys",
  "Wendys", "Subway", "Wendys", "Subway", "Subway", "Subway", "Subway",
  "Subway", "Subway")
```

Because of compiling Rmd file, I put these data by string vector instead of using the 'scan' function.
If we use the 'scan' function, we can see below block from R iterpretator.


>\> restaurant <- scan(what="character")
>
>1: Wendys McDonalds Subway Subway Subway Wendys
>
>7: Wendys Subway Wendys Subway Subway Subway
>
>13: Subway Subway Subway
>
>16: 
>
>Read 15 items
>
>\> restaurant
>
> [1] "Wendys"    "McDonalds" "Subway"    "Subway"    "Subway"    "Wendys"   
>
> [7] "Wendys"    "Subway"    "Wendys"    "Subway"    "Subway"    "Subway"
>
>[13] "Subway"    "Subway"    "Subway"
>


**b. Use the table function to find the frequencies of students who prefer the three restaurants.**

```{r}
table(restaurant)
```


**c. Compute the proportions of students in each category.**

```{r}
restaurant_factor <- factor(restaurant)
restaurant_prop_table <- table(restaurant_factor) / length(restaurant_factor)
restaurant_prop_table
```

**d. Construct two different graphical displays of the proportions.**

```{r}
plot(restaurant_prop_table,
  xlab = "restaurant", ylab = "proportion", main = "prefered restaurant")
barplot(restaurant_prop_table,
  xlab = "restaurant", ylab = "proportion", main = "prefered restaurant")
```


\vspace{1cm}


### \noindent Problem2.
**(Albert and Rizzo Chapter 3, problem 3.)**

**(Does baseball hitting data follow a binomial distribution?).**
**Albert Pujols is a baseball player who has $n$ opportunities to hit in a single**
**game. If $y$ denotes the number of hits for a game, then it is reasonable to**
**assume that $y$ has a binomial distribution with sample size $n$ and probability**
**of success $p = 0.312$, where $0.312$ is Pujols’ batting average (success rate) for**
**the 2010 baseball season.**

**a. In 70 games Pujols had exactly $n = 4$ opportunities to hit and the number of hits $y$ in these 70 games is tabulated in the following table.**

|Number of hits | Frequency|
|---|----|
| 0 | 17 |
| 1 | 31 |
| 2 | 17 |
|3 or more | 5 |

Use the dbinom function to compute the expected counts and the chisq.test
function to test if the counts follow a binomial(4, 0.312) distribution.

```{r}
observed_hit <- c(17, 31, 17, 5)
names(observed_hit) <- c("0", "1", "2", ">=3")
expected_prob <- dbinom(0:4, size = 4, prob = 0.312)
expected_hit <- expected_prob * 70
collapsed_expected_hit <- c(expected_hit[1:3], sum(expected_hit[4:5]))
collapsed_expected_prob <- c(expected_prob[1:3], sum(expected_prob[4:5]))
chisq_test_table <- data.frame(observed_hit, collapsed_expected_hit)
chisq_test_table

plot(0:3, collapsed_expected_hit, type = "h", lwd = 2, lty = 1, ylab = "Count",
  xlim = c(0, 3.5), ylim = c(0, 32), xlab = "Category", xaxt = "n")
axis(1, c(0, 1, 2, 3))
lines(0:3 + .1, observed_hit, type = "h", lwd = 2, lty = 2)
legend("topleft", legend = c("expected", "observed"),
  lty = c(1, 2), lwd = c(2, 2))


chisq_statistic <- sum((observed_hit - collapsed_expected_hit)^2 / collapsed_expected_hit)
p_val <- 1 - pchisq(chisq_statistic, length(observed_hit) - 1)

c(chisq_statistic, p_val)

# R built-in function
chisq.test(observed_hit, p = collapsed_expected_prob)
```

The hypotheses are:

H0: the observed 'number of hits' data follow a binomial(4, 0.312) distribution.

H1: Not H0.

The chi-square value is 0.97692, and it follows chi-square distribution with degree of freedom 3 under H0.
The p value is 0.8068.
Under a significance level less than 0.8068 (like 0.05 or 0.01), we cannot reject the H0.
Thus, we cannot say that the data do not follow binomial(4, 0.312) distribution.

```{r}
chisq.test(observed_hit, p = collapsed_expected_prob)$residual
```
All residual values are close to 0.


**b. In 25 games Pujols had exactly $n = 5$ opportunities to hit and the number**
**of hits y in these 25 games is shown in the table below.**

|Number of hits | Frequency|
|---|----|
| 0 | 5 |
| 1 | 5 |
| 2 | 4 |
|3 or more | 11 |

Use the chisq.test function to test if the counts follow a binomial(5, 0.312) distribution.


```{r}
observed_hit5 <- c(5, 5, 4, 11)
names(observed_hit5) <- c("0", "1", "2", ">=3")
expected_prob5 <- dbinom(0:5, size = 5, prob = 0.312)
expected_hit5 <- expected_prob5 * 25

collapsed_expected_hit5 <- c(expected_hit5[1:3], sum(expected_hit5[4:6]))
collapsed_expected_prob5 <- c(expected_prob5[1:3], sum(expected_prob5[4:6]))
chisq_test_table5 <- data.frame(observed_hit5, collapsed_expected_hit5)
chisq_test_table5

plot(0:3, collapsed_expected_hit5, type = "h", lwd = 2, lty = 1,
  ylab = "Count", xlim = c(0, 3.5), ylim = c(0, 15),
  xlab = "Category", xaxt = "n")
axis(1, c(0, 1, 2, 3))
lines(0:3 + .1, observed_hit5, type = "h", lwd = 2, lty = 2)
legend("topleft", legend = c("expected", "observed"),
  lty = c(1, 2), lwd = c(2, 2))


chisq.test(observed_hit5, p = collapsed_expected_prob5)
```

The hypotheses are:

H0: the observed 'number of hits' data follow a binomial(5, 0.312) distribution.

H1: Not H0.

The chi-square statistic value is 13.359, and it follows chi-square distribution with degree of freedom 3 under H0.
The p value is 0.003922.
Under a significance level larger than 0.003922 (like 0.05 or 0.01), we can reject the H0.
Thus, we can say that the data do not follow binomial(5, 0.312) distribution.

However, Let's take a look about the observed values and expected values table.
The expected value for '>=3' case is less than 5.
Thus, after collapsing the cagetory, let's repeat the chi-square test procedure.

```{r}
chisq.test(observed_hit5, p = collapsed_expected_prob5)$residual
```

The residual of class '>=3' is too large to be considered as coming from the standard normal distribution.


```{r}
collapsed2_observed_hit5 <- c(5, 5, 4 + 11)
names(observed_hit5) <- c("0", "1", ">=2")
collapsed2_expected_hit5 <- c(expected_hit5[1:2], sum(expected_hit5[3:6]))
collapsed2_expected_prob5 <- c(expected_prob5[1:2], sum(expected_prob5[3:6]))
collapsed2_chisq_test_table5 <- data.frame(collapsed2_observed_hit5, collapsed2_expected_hit5)
collapsed2_chisq_test_table5

plot(0:2, collapsed2_expected_hit5, type = "h", lwd = 2, lty = 1,
  ylab = "Count", xlim = c(0, 2.5), ylim = c(0, 18),
  xlab = "Category", xaxt = "n")
axis(1, 0:2)
lines(0:2 + .1, collapsed2_observed_hit5, type = "h", lwd = 2, lty = 2)
legend("topleft", legend = c("expected", "observed"),
  lty = c(1, 2), lwd = c(2, 2))

# R built-in function
chisq.test(collapsed2_observed_hit5, p = collapsed2_expected_prob5)
```

Again, the hypotheses are

H0: the observed 'number of hits' data follow a binomial(5, 0.312) distribution.

H1: Not H0.

The chi-square value is 2.4815, and it follows chi-square distribution with degree of freedom 2 under H0.
The p value is 0.2892
Under a significance level less than 0.2892 (like 0.05 or 0.01), we cannot reject the H0.
Thus, we cannot say that the data do not follow binomial(5, 0.312) distribution.

Here are residuals.
```{r}
chisq.test(collapsed2_observed_hit5, p = collapsed2_expected_prob5)$residual
```
All values are close to 0.

Note that test results have varied by how we collapse the categories.


\vspace{1cm}


### \noindent Problem3.
**(Albert and Rizzo Chapter 3, problem 4.)**
3.4 (Categorizing ages in the twins dataset). The variable AGE gives the age (in years) of twin 1.

```{r}
twins <- read.table("Rx-data/twins.dat.txt",
  header = TRUE, sep = ",", na.strings = ".")
names(twins)
head(twins)
```

a. Use the cut function on AGE with the breakpoints 30, 40, and 50 to create
a categorized version of the twin’s age.

```{r}
categorized_age <- cut(twins$AGE, breaks = c(0, 30, 40, 50, max(twins$AGE)))
```

b. Use the table function to find the frequencies in the four age categories.
```{r}
tb_categorized_age <- table(categorized_age)
```

c. Construct a graph of the proportions in the four age categories.

```{r}
proptb_categorized_age <- prop.table(tb_categorized_age)
barplot(proptb_categorized_age)
```


\vspace{1cm}


### \noindent Problem4.
**(Albert and Rizzo Chapter 3, problem 5.)**
**3.5 (Relating age and wage in the twins dataset). The variables AGE**
**and HRWAGEL contain the age (in years) and hourly wage (in dollars) of twin 1.**

**a. Using two applications of the cut function, create a categorized version**
**of AGE using the breakpoints 30, 40, and 50, and a categorized version of**
**HRWAGEL using the same breakpoints as in Section 3.3.**

```{r}
# twins$AGE
# twins$HRWAGEL
categorized_age <- cut(twins$AGE, breaks = c(0, 30, 40, 50, max(twins$AGE)))
categorized_hrwageL <- cut(twins$HRWAGEL, breaks = c(0, 7, 13, 20, 150))
```


**b. Using the categorized versions of AGE and HRWAGEL, construct a contingency table of the two variables using the function table.**

```{r}
table(categorized_age, categorized_hrwageL)
```

**c. Use the prop.table function to find the proportions of twins in each age**
**class that have the different wage groups.**

```{r}
twin_CT <- prop.table(table(categorized_age, categorized_hrwageL), margin = 1)
twin_CT
```

**d. Construct a suitable graph to show how the wage distribution depends on**
**the age of the twin.**

Here is the mosaic plot.
```{r}
plot(prop.table(table(categorized_age, categorized_hrwageL)),
  ylab = "hourly wage ($)", xlab = "age (year)", main = "twin")
```

```{r}
barplot(t(twin_CT), ylim = c(0, 1.3), ylab = "hourly wage: PROPORTION",
  legend.text = dimnames(twin_CT)$categorized_hrwageL,
  args.legend = list(x = "top"), xlab = "age")
```
```{r}
barplot(t(twin_CT), beside = T,
  legend.text = dimnames(twin_CT)$categorized_hrwageL,
  args.legend = list(x = "topright"),
  ylab = "hourly wage: PROPORTION", xlab = "age")
```

**e. Use the conditional proportions in part (c) and the graph in part (d) to**
**explain the relationship between age and wage of the twins.**

Depending on the age, the hourly wage varies.

In the younger generation under the age of 30, nobody could earn more than 20 dollars in an hour.
About 45% of 'twin 1' people in this generation class make less or equal than \$ 7 in an hour.
And, the number of 'twin 1' people in this generation decrease as the wage increase.

About 43% of 'twin 1' people who are between 30 and 40 years old make between $7 and $13 in an hour.
Comparing to the younger generation, the ratio of people earning too low a wage has decreased.
Not only that, the ratios of people earning $13~20 and $20~150 in an hour have increased.

'Twin 1' people who are between 40 and 50 years old are earning more money than younger 'twin 1' people.
About 32% are earning more than $20 in an hour. The other three wage classes have the same proportion.

However, in the case of 'twin 1' people in the oldest group, the wage level decreases again.
About 35% of 'twin 1' people in this group is earning between $7 and $13 in an hour.

To sum up, the general hourly wage level tends to increase until 50 years old, and then decrease.


\vspace{1cm}


### \noindent Problem5.
**(Albert and Rizzo Chapter 3, problem 6.)**
**3.6 (Relating age and wage in the twins dataset, continued).**

**a. Using the contingency table of the categorized version of AGE and HRWAGEL**
**and the function chisq.test, perform a test of independence of age and wage.**
**Based on this test, is there significant evidence to conclude that age and wage are dependent?**

H0: AGE and HRWAGEL are independent

H1: AGE and HRWAGEL are not independent

```{r}
twins_chisq <- chisq.test(categorized_age, categorized_hrwageL)
twins_chisq$expected
twins_chisq$observed
twins_chisq

```

The chi-square statistic value is 24.771.
Under H0, the value follows chi-square distribution with d.f. 9.
The p-value is 0.003235, so we can reject the H0 if we use higher significance level than 0.003235 (like 0.01 or 0.05).
Thus, we can conclude that the two variables are not independent.

**b. Compute and display the Pearson residuals from the test of independence.**
**Find the residuals that exceed 2 in absolute value.**

```{r}
twins_chisq$residuals
which(abs(twins_chisq$residuals) > 2)
```

There are two residuals that exceed 2 in absolute value.
One is age (0,30] with hourly wage (20, 150].
The other one is age (40,50] with hourly wage (20, 150].

**c. Use the function mosaicplot with the argument shade=TRUE to construct**
**a mosaic plot of the table counts showing the extreme residuals.**

```{r}
plot(table(categorized_age, categorized_hrwageL), shade = TRUE)
# ?plot.table
```
Because the class which is age (0,30] with hourly wage (20, 150] has no proportion,
the class is not visible.

**d. Use the numerical and graphical work from parts (b) and (c) to explain**
**how the table of age and wages differs from an independence structure.**

If age and wages were independent, the chi-square test would not be rejected.
Not only that, the size of squares in the mosaic plot would be similar.
However, as we saw at the test result and the realized mosaic plot, it shows a different pattern from the independent case.


\vspace{1cm}


### \noindent Problem6.
**(Albert and Rizzo Chapter 3, problem 8.)**
**3.8 (Are the digits of $\pi$ random?). The National Institute of Standards**
**and Technology has a web page that lists the first 5000 digits of the irrational**
**number $\pi$. One can read these digits into R by means of the script**


```{r}
pidigits <- read.table(
    "http://www.itl.nist.gov/div898/strd/univ/data/PiDigits.dat",
     skip = 60)
head(pidigits)
```

**a. Use the table function to construct a frequency table of the digits 1 through 9.**
```{r}
freq_pidigits <- table(pidigits)
freq_pidigits
```


**b. Construct a bar plot of the frequencies found in part (a).**

```{r}
prop_pidigits <- prop.table(freq_pidigits)
barplot(prop_pidigits, main = "pi's first 5000 digits: proportion")
```


**c. Use the chi-square test, as implemented in the chisq.test function, to**
**test the hypothesis that the digits 1 through 9 are equally probable in the**
**digits of $\pi$.**

Let $p_i$ be a probability of $i$ at $\pi$'s digits

H0: digits are equally probable ($p_0 = p_1 = ... = p_9 = 0.1$)

H1: digits are not equally probable ($p_0 = p_1 = ... = p_9 != 0.1$)

```{r}
chisq.test(freq_pidigits, p = rep(0.1, 10))
```

The p-value is 0.3224. We cannot reject H0 if we use a significance level less than 0.3224 (like 0.05 or 0.01.)
Hence, we cannot say that each digit is not equally probable.


\vspace{1cm}


### \noindent Problem7.
**(Albert and Rizzo Chapter 5, problem 2.)**

**5.2 (Relationship between the percentages of small classes and large classes).**
**The variables Pct.20 and Pct.50 in the college dataset contain**
**respectively the percentage of “small classes” (defined as 20 or fewer students)**
**and the percentage of “large classes” (defined as 50 or more students) in the National Universities.**

```{r}
college_dat <- read.table("Rx-data/college.txt", header = TRUE, sep = "\t")
college_dat <- subset(college_dat, complete.cases(college_dat))
names(college_dat)
```

**a. Use the plot function to construct a scatterplot of Pct.20 (horizontal)**
**against Pct.50 (vertical).**
```{r}
plot(college_dat$Pct.20, college_dat$Pct.50,
  pch = 19, xlab = "Pct.20", ylab = "Pct.50")
```

**b. Use the line function to find a resistant line to these data. Add this**
**resistant line to the scatterplot constructed in part a.**
```{r}
college_resistant_line_fit <- line(college_dat$Pct.20, college_dat$Pct.50)
plot(college_dat$Pct.20, college_dat$Pct.50,
  pch = 19, xlab = "Pct.20", ylab = "Pct.50")
abline(college_resistant_line_fit, col = "red")
```

**c. If 60% of the classes at a particular college have 20 or fewer students, use**
**the fitted line to predict the percentage of classes that have 50 or more students.**

```{r}
college_resistant_coeff <- coef(college_resistant_line_fit)
predicted_percentage_pct50 <- 60 * college_resistant_coeff[2] + college_resistant_coeff[1]
print(predicted_percentage_pct50)
```

According to the prediction result, about 7.8 percent of classes have 50 or more students.


**d. Construct a graph of the residuals (vertical) against Pct.20 (horizontal)**
**and add a horizontal line at zero (using the abline function).**
```{r}
college_resistant_resid <- college_resistant_line_fit$residuals
plot(college_dat$Pct.20, college_resistant_resid)
abline(h = 0)
```

**e. Is there a distinctive pattern to the residuals? (Compare the sizes of the**
**residuals for small Pct.20 and the sizes of the residuals for large Pct.50.)**


```{r}
plot(college_dat$Pct.20, college_resistant_resid)
abline(h = 0)
outlier15_idx <- which(abs(resid) > 15)
college_dat[outlier15_idx, ]
college_resistant_resid[outlier15_idx]
text(college_dat$Pct.20[outlier15_idx], college_resistant_resid[outlier15_idx],
  labels = college_dat[outlier15_idx, 1], pos = 1)
```

In the residual plot in (d), we can find some outliers.
UC San Diego's residual is about 17.7 which is huge relative to other colleges. 

Next, let's draw the residual plot with Pct.50 values.

```{r}
plot(college_dat$Pct.50, resid)
abline(h = 0)
text(college_dat$Pct.50[outlier15_idx], college_resistant_resid[outlier15_idx],
  labels = college_dat[outlier15_idx, 1], pos = 2)
```

An existance of a pattern in a residual plot shows that the fit wasn't that good.
However, clearly, an increasing pattern is visible in this residual plot!


**f. Use the identify function to identify the schools that have residuals that**
**exceed 10 in absolute value. Interpret these large residuals in the context**
**of the problem.**

```{r}
# plot(dat$Pct.50, resid)
# identify(dat$Pct.50, resid, n = 6, labels = dat$School)
```

The 'identify' function does not work in the R-markdown.
(we have to click 6 points after the line,
but we cannot do that during the r-markdown compile process.)

So, I will make the same output in another way.
```{r}
plot(college_dat$Pct.50, college_resistant_resid)
abline(h = 0)
abline(h = 10, lty = 2)
abline(h = -10, lty = 2)
outlier10upper_idx <- which(college_resistant_resid > 10)
outlier10lower_idx <- which(college_resistant_resid < -10)
outlier10upper_school <- college_dat$School[outlier10upper_idx]
outlier10lower_school <- college_dat$School[outlier10lower_idx]
outlier10upper_school
outlier10lower_school[3]
for (i in seq_len(length(outlier10upper_idx))) {
  text(college_dat$Pct.50[outlier10upper_idx[i]],
    college_resistant_resid[outlier10upper_idx[i]],
    labels = outlier10upper_school[i], pos = 2)
}
for (i in seq_len(length(outlier10lower_idx))) {
  text(college_dat$Pct.50[outlier10lower_idx[i]],
    college_resistant_resid[outlier10lower_idx[i]],
    labels = outlier10lower_school[i], pos = 4)
}
```


\vspace{1cm}


### \noindent Problem8.
**(Albert and Rizzo Chapter 5, problem 3.)**

**5.3 (Relationship between acceptance rate and “top-ten” percentage).**
**The variables Accept.rate and Top.10 in the college dataset contain**
**respectively the acceptance rate and the percentage of incoming students in**
**the top 10 percent of their high school class in the National Universities.**
**One would believe that these two variables are strongly associated, since, for**
**example, "exclusive" colleges with small acceptance rates would be expected**
**to have a large percentage of "top-ten" students.**

```{r}
names(college_dat)

```


**a. Explore the relationship between Accept.rate and Top.10.**
**This exploration should include a graph and linear fit that describe the basic pattern**
**in the relationship and a residual graph that shows how schools differ from**
**the basic pattern.**

```{r}
par(mfrow = c(1, 2))
plot(college_dat$Top.10, college_dat$Accept.rate)
college_t10_ar_resistant_line_fit <- line(college_dat$Top.10, college_dat$Accept.rate)
coef(college_t10_ar_resistant_line_fit)
abline(college_t10_ar_resistant_line_fit, col = "red")

college_t10_ar_resid <- college_t10_ar_resistant_line_fit$residuals
plot(college_dat$Top.10, college_t10_ar_resid, ylim = c(-32, 52))
abline(h = 0)
abline(h = c(-30, 30), lty = 2)

```

Bayesically, 'Top.10' are negatively correlated with 'Accept.rate'.
The slope of Tukey's resistant line is -0.575. So, students who were in the top ten
percent of their high school class is more likely to enter college that have low acceptance rate.

Most residuals are between -30 and 30.
However, we can find two exceptional residual values in the upper-right hand corner of the residual plot.


**b. Schools are often classified into "elite" and "non-elite" colleges depending**
**on the type of students they admit. Based on your work in part a, is there**
**any evidence from Accept.rate and Top.10 that schools do indeed cluster**
**into "elite" and "non-elite" groups? Explain.**

Let me add an auxiliary line at Top.10 = 60.

```{r}
par(mfrow = c(1, 2))
plot(college_dat$Top.10, college_dat$Accept.rate)
abline(college_t10_ar_resistant_line_fit, col = "red")
abline(v = 60, lty = 2, col = "blue")

plot(college_dat$Top.10, college_t10_ar_resid, ylim = c(-32, 52))
abline(h = 0)
abline(h = c(-30, 30), lty = 2)
abline(v = 60, lty = 2, col = "blue")
```

Yes. If we compare the left part with the right part, we can try creating a cluster.
Most students in non-elite group entered colleges which have high acceptance rates.
And, the points corresponding to the non-elite students are pretty well put together.

In contrast, students in elite group tend to go to low-acceptance-rate collages.
However, in the data points of elite group, acceptance rates have great volatility, with some outliers.

To see it more clearly, I'm going to draw two more pictures: the stripchart and the boxplot of our data.

```{r}
par(mfrow = c(1, 2))
college_dat$elite <- college_dat$Top.10 > 60
stripchart(Accept.rate ~ elite, method = "stack", pch = 19,
  xlab = "Acceptance rate",
  ylab = "Elite", data = college_dat)

boxplot(Accept.rate ~ elite, data = college_dat, horizontal = TRUE)
```

Two boxes in the boxplot do not overlap at all.
Two median value are quite far.


\vspace{1cm}


### \noindent Problem9.
**(Albert and Rizzo Chapter 5, problem 5.)**

**5.5 (Exploring percentages of full-time faculty).**
**The variable Full.time in the college dataset (see Example 5.3) contains the percentage of faculty**
**who are hired full-time in the group of National Universities.**

**a. Using the hist function, construct a histogram of the full-time percentages**
**and comment on the shape of the distribution.**

**b. Use the froot and flog transformations to reexpress the full-time percentages.**
**Construct histograms of the collection of froots and the collection of flogs.**
**Is either transformation successful in making the full-time percentages approximately symmetric?**

**c. For data that is approximately normally distributed, about 68% of the**
**data fall within one standard deviation of the mean. Assuming you have**
**found a transformation in part (b) that makes the full-time percentages**
**approximately normal, find an interval that contains roughly 68% of the**
**data on the new scale.**


\vspace{1cm}


### \noindent Problem10.
**(Albert and Rizzo Chapter 5, problem 6.)**
**5.6 (Exploring alumni giving rates). The variable Alumni.giving contains**
**the percentage of alumni from the college who make financial contributions.**
**a. Construct a “stacked” dotplot of the alumni giving percentages using the**
**stripchart function.**
**b. Identify the names of the three schools with unusually large giving percentages.**
**c. It can be difficult to summarize these giving percentages since the distribution is right-skewed.**
**One can make the dataset more symmetric by**
**applying either a square root transformation or a log transformation.**

roots = sqrt(college$Alumni.giving)
logs = log(college$Alumni.giving)

**Apply both square root and log transformations. Which transformation**
**makes the alumni giving rates approximately symmetric?**


\vspace{1cm}


### \noindent Problem11.
**(Albert and Rizzo Chapter 5, problem 7.)**

**5.7 (Exploring alumni giving rates (continued)).**
**In this exercise, we focus on the comparison of the alumni giving percentages**
**between the four tiers of colleges.**



**a. Using the stripchart function with the stacked option, construct parallel**
**dotplots of alumni giving by tier.**
**b. As one moves from Tier 4 to Tier 1, how does the average giving change?**
**c. As one moves from Tier 4 to Tier 1, how does the spread of the giving**
**rates change?**
**d. We note from parts (b) and (c), that small giving rates tend to have**
**small variation, and large giving rates tend to have large variation. One**
**way of removing the dependence of average with spread is to apply a**
**power transformation such as a square root or a log. Construct parallel**
**stripcharts of the square roots of the giving rates, and parallel boxplots of**
**the log giving rates.**
**e. Looking at the two sets of parallel stripcharts in part (d), were the square**
**root rates or the log rates successful in making the spreads approximately**
**the same between groups?**
